---
title: "20Newsgroup Analysis"
author: "Grace Lawley"
output: 
  html_document:
    theme: 'flatly'
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(stringr)
library(heplots) # boxM()
library(effectsize) # eta_squared()
set.seed(1234)


lda20_gamma_ilr <- readRDS("data/lda20_gamma_ilr.RDS")
lda20_gamma_ilr <- readRDS("data/lda20_gamma_ilr.RDS")
```

LDA model was fit using $k=20$ on documents from 4 labels: 

1. `comp.sys.ibm.pc.hardware`
1. `comp.sys.mac.hardware`
1. `rec.sport.baseball`
1. `rec.sport.hockey`


```{r}
lda20_gamma_ilr %>% 
  count(label)
```


# Between broader categories (x1)

## (1) `comp.sys.*` vs. `rec.sport.*`

> Compare topic distributions between documents from `comp.sys.[ibm.pc|mac].hardware` vs. documents from `rec.sport.[baseball|hockey]`.


```{r}
# prepare dataframe
comp_v_rec <- lda20_gamma_ilr %>% 
  mutate(label = str_extract(label, "[a-z]+"))

comp_v_rec %>% 
  count(label)
```


```{r}
# reshape dataframe into matrices
comp_v_rec_X <- comp_v_rec %>%
  select(-label) %>% 
  as.matrix()
comp_v_rec_Y <- comp_v_rec$label
```



### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(comp_v_rec_X, comp_v_rec_Y)
```

### MANOVA

```{r}
comp_v_rec_manova <- manova(comp_v_rec_X ~ comp_v_rec_Y, data = comp_v_rec)
summary(comp_v_rec_manova)
```


### Partial eta-squared (effect size)

```{r}
eta_squared(comp_v_rec_manova)
```



# Between subcategories (x2)

## (1) `comp.sys.ibm.pc.hardware` vs. `comp.sys.mac.hardware`

> Compare topic distributions between documents from `comp.sys.ibm.pc.hardware` vs. documents from `comp.sys.mac.hardware`.



```{r}
# prepare data
mac_v_pc <- lda20_gamma_ilr %>%
  filter(str_detect(label, "comp.sys"))

mac_v_pc %>% 
  count(label)
```

```{r}
# reshape dataframe into matrices
mac_v_pc_X <- mac_v_pc %>%
  select(-label) %>% 
  as.matrix() 
mac_v_pc_Y <- mac_v_pc$label
```

### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(mac_v_pc_X, mac_v_pc_Y)
```

### MANOVA

```{r}
mac_v_pc_manova <- manova(mac_v_pc_X ~ mac_v_pc_Y, data = mac_v_pc)
summary(mac_v_pc_manova)
```


### Partial eta-squared (effect size)

```{r}
eta_squared(mac_v_pc_manova)
```


## (2) `rec.sport.baseball` vs. `rec.sport.hockey`

> Compare topic distributions between documents from `rec.sport.baseball` and documents from `rec.sport.hockey`.

```{r}
# prepare data
baseball_v_hockey <- lda20_gamma_ilr %>%
  filter(str_detect(label, "rec.sport"))

baseball_v_hockey %>% 
  count(label)
```

```{r}
# reshape dataframe into matrices
baseball_v_hockey_X <- baseball_v_hockey %>%
  select(-label) %>% 
  as.matrix() 
baseball_v_hockey_Y <- baseball_v_hockey$label
```

### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(baseball_v_hockey_X, baseball_v_hockey_Y)
```

### MANOVA

```{r}
baseball_v_hockey_manova <- manova(baseball_v_hockey_X ~ baseball_v_hockey_Y, data = baseball_v_hockey)
summary(baseball_v_hockey_manova)
```


### Partial eta-squared (effect size)

```{r}
eta_squared(baseball_v_hockey_manova)
```


# Within a single topic (x4)

## (1) Within `comp.sys.ibm.pc.hardware`

> Randomly split documents from `comp.sys.ibm.pc.hardware` into two groups. Compare topic distributions between documents from `comp.sys.ibm.pc.hardware_1` and documents from `comp.sys.ibm.pc.hardware_2`.

```{r}
# prepare data
pc_v_pc <- lda20_gamma_ilr %>% 
  filter(str_detect(label, "comp.sys.ibm.pc.hardware")) %>%
  # randomly split into 2 groups
  mutate(x = sample(c(1, 2), nrow(.), replace = TRUE)) %>% 
  mutate(label = paste0(label, "_", x)) %>% 
  select(-x) %>% 
  mutate(label = as.factor(label))

pc_v_pc %>% 
  count(label)
```

```{r}
# reshape dataframe into matrices
pc_v_pc_X <- pc_v_pc %>%
  select(-label) %>% 
  as.matrix()
pc_v_pc_Y <- pc_v_pc$label
```

### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(pc_v_pc_X, pc_v_pc_Y)
```

### MANOVA

```{r}
pc_v_pc_manova <- manova(pc_v_pc_X ~ pc_v_pc_Y, data = pc_v_pc)
summary(pc_v_pc_manova)
```

### Partial eta-squared (effect size)

```{r}
eta_squared(pc_v_pc_manova)
```

## (2) Within `comp.sys.mac.hardware`

> Randomly split documents from `comp.sys.mac.hardware` into two groups. Compare topic distributions between documents from `comp.sys.mac.hardware_1` and documents from `comp.sys.mac.hardware_2`.


```{r}
# prepare data
mac_v_mac <- lda20_gamma_ilr %>% 
  filter(str_detect(label, "comp.sys.mac.hardware")) %>%
  # randomly split into 2 groups
  mutate(x = sample(c(1, 2), nrow(.), replace = TRUE)) %>% 
  mutate(label = paste0(label, "_", x)) %>% 
  select(-x) %>% 
  mutate(label = as.factor(label))

mac_v_mac %>% 
  count(label)
```

```{r}
# reshape dataframe into matrices
mac_v_mac_X <- mac_v_mac %>%
  select(-label) %>% 
  as.matrix()
mac_v_mac_Y <- mac_v_mac$label
```

### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(mac_v_mac_X, mac_v_mac_Y)
```

### MANOVA

```{r}
mac_v_mac_manova <- manova(mac_v_mac_X ~ mac_v_mac_Y, data = mac_v_mac)
summary(mac_v_mac_manova)
```

### Partial eta-squared (effect size)

```{r}
eta_squared(mac_v_mac_manova)
```


## (3) Within `rec.sport.baseball`

> Randomly split documents from `rec.sport.baseball` into two groups.  Compare topic distributions between documents from `rec.sport.baseball_1` and documents from `rec.sport.baseball_2`.

```{r}
# prepare data
baseball_v_baseball <- lda20_gamma_ilr %>% 
  filter(str_detect(label, "rec.sport.baseball")) %>%
  # randomly split into 2 groups
  mutate(x = sample(c(1, 2), nrow(.), replace = TRUE)) %>% 
  mutate(label = paste0(label, "_", x)) %>% 
  select(-x)

baseball_v_baseball %>% 
  count(label)
```

```{r}
# reshape dataframe into matrices
baseball_v_baseball_X <- baseball_v_baseball %>%
  select(-label) %>% 
  as.matrix()
baseball_v_baseball_Y <- baseball_v_baseball$label
```

### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(baseball_v_baseball_X, baseball_v_baseball_Y)
```

### MANOVA

```{r}
baseball_v_baseball_manova <- manova(baseball_v_baseball_X ~ baseball_v_baseball_Y, data = baseball_v_baseball)
summary(baseball_v_baseball_manova)
```

### Partial eta-squared (effect size)

```{r}
eta_squared(baseball_v_baseball_manova)
```

## (4) Within `rec.sport.hockey`

> Randomly split documents from `rec.sport.hockey` into two groups. Compare topic distributions between documents from `rec.sport.hockey_1` and documents from `rec.sport.hockey_2`.

```{r}
# prepare data
hockey_v_hockey <- lda20_gamma_ilr %>% 
  filter(str_detect(label, "rec.sport.hockey")) %>%
  # randomly split into 2 groups
  mutate(x = sample(c(1, 2), nrow(.), replace = TRUE)) %>% 
  mutate(label = paste0(label, "_", x)) %>% 
  select(-x) %>% 
  mutate(label = as.factor(label))

hockey_v_hockey %>% 
  count(label)
```

```{r}
# reshape dataframe into matrices
hockey_v_hockey_X <- hockey_v_hockey %>%
  select(-label) %>% 
  as.matrix()
hockey_v_hockey_Y <- hockey_v_hockey$label
```

### Box’s M Test for homogeneity of covariance matrices

If $p < 0.001$, this MANOVA assumption not met (but okay to proceed)

```{r}
boxM(hockey_v_hockey_X, hockey_v_hockey_Y)
```

### MANOVA

```{r}
hockey_v_hockey_manova <- manova(hockey_v_hockey_X ~ hockey_v_hockey_Y, data = hockey_v_hockey)
summary(hockey_v_hockey_manova)
```

### Partial eta-squared (effect size)

```{r}
eta_squared(hockey_v_hockey_manova)
```

